---
title: "Corlla Regression"
output:
word_document: default
pdf_document: default
html_document: default
date: "May 11, 2019"
---

```{r, include=FALSE}
library(stats)
library(mosaic)
library(tidyverse)
library(DT)
library(plotly)
library(pander)
library(car)
corp <- read.csv("../Data/corp.csv", header=TRUE)
```

## Background

I have no idea why my data variables don't work in either app or simple markdown file so I apologize for the text file.

## Analysis

Given a R^2 of .7279, I was surprised at the somewhat strong trend of Pricing based on Mileage.
Plotting my regression led me to think that I needed to trasnform it based on the assumptions(*listed at the bottom*).
I would say that there is constant variance and some independent error terms but not much linearity or normal error terms.

```{r}
corplm<-lm(Price~Mileage,data=corp)
plot(Price~Mileage,data=corp)
abline(corplm)
summary(corplm)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
summary(corplm)
```

## Assumptions

```{r}
plot(corplm)
```