---
title: "Different Models"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

```{r, include=FALSE}

library(ggplot2)
library(car)
library(tidyverse)
library(Ecdat)
library(pander)

set.seed(101) #Allows us to always get the same "random" sample
#Change to a new number to get a new sample

n <- 30 #set the sample sizes
```

## Background

I've created three different types of regression models with arbitrary parameters of my choosing which I used to generate data around each model. The goa is to create an estimate for each model parameter, and see how close the estimate gets to the 'true' model. 

To do this I chose to plot the true regression using a solid line, and the estimated regression using the dashed line. I also thought it might be interesting to plot a simple linear regression line using dashed gray line and compare the r-squared values with the more complex models to show how proportion of variance explained differs between them.

## Quadratic Model

---

$$
  Y_i = \beta_0 + \beta_1 X_i + \beta_2X_i^2 + \epsilon_i
$$

---

The true regression model used to generate the data is 

$E\{Y_i\} = -1000 + 70X_i - 1X_i^2$, represented by the solid blue line.

This was estimated to be 

$\hat{Y_i} = -872.9173 + 60.467X_i - .843X_i^2$, shown using the dotted blue line.


```{r}
X_i <- runif(n, 15, 45)
#Gives n random values from a uniform distribution between 15 to 45.

beta0 <- -1000 #Our choice for the y-intercept. 
beta1 <- 70 #Our choice for the slope. 
beta2 <- -1
sigma <- 35 #Our choice for the std. deviation of the error terms.


epsilon_i <- rnorm(n, 0, sigma)
#Gives n random values from a normal distribution with mean = 0, st. dev. = sigma.

Y_i <- beta0 + beta1*X_i + beta2*X_i^2 + epsilon_i
#Create Y using the normal error regression model

fabData <- data.frame(y=Y_i, x=X_i)

#In the real world, we begin with data (like fabData) and try to recover the model that 
# (we assume) was used to created it.

fab.lm <- lm(y ~ x + I(x^2), data=fabData) #Fit an estimated regression model to the fabData.
simple.lm <- lm(y ~ x, data=fabData)

b <- coef(fab.lm)

ggplot(fabData) +
  geom_point(aes(y = y, x = x)) +
  stat_function(fun = function(x) beta0 + beta1 * x + beta2 * x^2, color = "skyblue") +
  stat_function(fun = function(x) b[1] + b[2] * x + b[3] * x^2, linetype="dashed", color="skyblue") +
  geom_abline(intercept = simple.lm$coefficients[1], slope = simple.lm$coefficients[2], color="gray", linetype="dashed") +
  labs(main = "test") +
  theme_minimal()

```

| Model   | $R^2$ | Adjusted $R^2$ |
|---------|-------|----------------|
| Simple  | `r summary(simple.lm)$r.squared` | `r summary(simple.lm)$adj.r.squared` |
| Quadratic   | `r summary(fab.lm)$r.squared`|   `r summary(fab.lm)$adj.r.squared`      |

## Two-lines Regression Model

---

$$
  Y_i = \beta_0 + \beta_1 X_1i + \beta_2X_2i + \beta_3 X_1i X_2i + \epsilon_i
$$

---

```{r}
beta0 <- 5 #Our choice for the y-intercept. 
beta1 <- 35 #Our choice for the slope. 
sigma <- 250 #Our choice for the std. deviation of the error terms.

beta0_2 <- 10 #Our choice for the y-intercept. 
beta1_2 <- 65 #Our choice for the slope. 
sigma_2 <- 150 #Our choice for the std. deviation of the error terms.

X_i <- runif(n, 15, 45)
X_i2 <- runif(n, 15, 45)

epsilon_i <- rnorm(n, 0, sigma) 
epsilon_i2 <- rnorm(n, 0, sigma_2)

Y_i <- beta0 + beta1*X_i + epsilon_i
Y_i2 <- beta0_2 + beta1_2*X_i2 + epsilon_i2

fabData <- data.frame(y=Y_i, x=X_i, type='one')
fabData <- rbind(fabData, data.frame(y=Y_i2, x=X_i2, type='two'))

# plot(y ~ x, data=fabData, col=c("skyblue", "orange")[as.factor(type)], pch=21, cex.main=1)

multi.lm <- lm(y ~ x + type + x:type, data=fabData)
simple.lm <- lm(y ~ x, data=fabData)

b <- coef(multi.lm)

```

The true model is $E\{Y_i\} = 5 + 35X_1i$, seen by the solid red line, and $\hat{Y_i} = (10 + 5) + (35 + 65)X_i$, seen by the solid blue line.

The estimated models are 

$\hat{Y_i} = `r b[1]` + `r b[2]`X_i$ (dashed red line). 

$\hat{Y_i} = (`r b[1]` + `r b[3]`) + (`r b[2]` + `r b[4]`)X_i$ (dashed blue line)

```{r}
ggplot(fabData, aes(y = y, x = x, color=as.factor(type))) +
  geom_point() +
  labs(color="type") +
  stat_function(fun = function(x) beta0 + beta1*x, color="red") +
  stat_function(fun = function(x) b[1] + b[2]*x, linetype="dashed", color="red") +
  stat_function(fun = function(x) beta0_2 + beta1_2*x) +
  stat_function(fun = function(x) (b[1] + b[3]) + (b[2] + b[4])*x, linetype="dashed") +
  geom_abline(intercept = simple.lm$coefficients[1], slope = simple.lm$coefficients[2], color="gray", linetype="dashed") +
  theme_minimal()
```

| Model   | $R^2$ | Adjusted $R^2$ |
|---------|-------|----------------|
| Simple  | `r summary(simple.lm)$r.squared` | `r summary(simple.lm)$adj.r.squared` |
| Two-lines   | `r summary(multi.lm)$r.squared`|   `r summary(multi.lm)$adj.r.squared` |

## Cubic Model

---

$$
  Y_i = \beta_0 + \beta_1 X_i + \beta_2X_i^2 + \beta_3X_i^3 + \epsilon_i
$$

---


```{r}
set.seed(101) #gives us the same randomness 
n <- 20 #sample size
X <- runif(n, -1.5, 2)

beta0 <- 1
beta1 <- -.1
beta2 <- 2.5
beta3 <- .5

Y <- beta0 + beta1*X + beta2*X^2 + beta3*X^3 + rnorm(n, 0, 1)


fake.lm <- lm(Y ~ X + I(X^2) + I(X^3), data = data.frame(Y=Y, X=X))
simple.lm <- lm(Y ~ X, data = data.frame(Y=Y, X=X))

b <- coef(fake.lm)
```
True model:

$E\{Y_i\}  = 1 - .1 X_i + 2.5X_i^2 + .5X_i^3 + \epsilon_i$ (solid black line).

Estimated model: 

$Y_i = `r b[1]` + `r b[2]` X_i + `r b[3]`X_i^2 + `r b[4]`X_i^3 + \epsilon_i$ (dashed black line).

```{r}
plot(Y ~ X, pch=21, col="lightgray", bg="steelblue", ylim=c(-5,22))

fabulousData <- data.frame(y = Y, x = X)

curve(b[1] + b[2] + b[3]*x^2 + b[4]*x^3, add = T)
curve(beta0 + beta1 + beta2*x^2 + beta3*x^3, add = T, lty = 2)
abline(simple.lm, lty=2, col="gray", lwd=.8)
```

| Model   | $R^2$ | Adjusted $R^2$ |
|---------|-------|----------------|
| Simple  | `r summary(simple.lm)$r.squared` | `r summary(simple.lm)$adj.r.squared` |
| Cubic   | `r summary(fake.lm)$r.squared`|   `r summary(fake.lm)$adj.r.squared` |

## Example with real dataset

I used the mtcars dataset to explain miles per gallon using horsepower. I applied a simple, quadratic, and cubic model. Both the cubic and quadratic model seemed to fit the data similarly, and they did a decent job, at least better than the simple regression.

```{r}

wage.cube.lm <- lm(mpg ~ hp + I(hp^2) + I(hp^3), data = mtcars)
wage.quad.lm <- lm(mpg ~ hp + I(hp^2), data = mtcars)
wage.simple.lm <- lm(mpg ~ hp, data = mtcars)
b <- coef(wage.cube.lm)
b2 <- coef(wage.quad.lm)
b3 <- coef(wage.simple.lm)

# curve(b[1] + b[2] * x + b[3] * x^2 + b[4] * x^3, add = T)
# curve(b2[1] + b2[2] * x + b2[3] * x^2, add = T)
# abline(wage.simple.lm)

ggplot(mtcars, aes(y= mpg, x = hp)) +
  geom_point() +
  stat_function(fun = function(x) b[1] + b[2] * x + b[3] * x^2 + b[4] * x^3, color="firebrick") +
  stat_function(fun = function(x) b2[1] + b2[2] * x + b2[3] * x^2, color="skyblue") +
  stat_function(fun = function(x) b3[1] + b3[2] * x, color="gray") +
  labs(title = "Quadratic and Cubic models perform about the same but better than simple regression model") +
  theme_bw()
```

<b>Quadratic model:</b>

$\hat{Y_i} = -40.21 - 0.2133X_i + .0004208X_i^2$

In the quadratic model the estimate of $\beta_0$, `r wage.quad.lm$coefficients[1]` is the y-intercept, the estimate for $\beta_1$, `r wage.quad.lm$coefficients[2]` will be used to determine where the vertex of the parabola is ($\frac{−\beta_1}{2\cdot\beta_2}$). The estimate for $\beta_2$, 0.0004208 determines which direction the parabola will face and how "steep" it is - larger values make for a steeper parabola.

<b>Cubic model:</b>

$\hat{Y_i} = 44.22 + -0.2945X_i + 0.0009115X_i^2 - 8.701086e-07X_i^3$

In this model, the 44.22 ($\beta_0$) estimate represents the Y-intercept as usual. -0.2945 ($\beta_1$) is not so easily interpreted but can be thought of as contributing to the location of the local minimum and maximum values. 0.0009115 ($\beta_2$) is similar to $\beta_1$. -8.701e-07 ($\beta_3$) determines whether the model is increasing or decreasing depending on it's sign.

<b>Two-line model:</b>

I'll use the same old dataset as everyone else for the two-line model, where an automatic transmission is represented as 0 so that it can cancel out part of the equation and a manual transmission is 1. The big twist is that I'm predicting qsec based on mpg.

```{r}
m.lm <- lm(qsec ~ mpg + am + mpg:am, data = mtcars)

b <- coef(m.lm)

ggplot(mtcars, aes(x = mpg, y = qsec, colour=as.factor(am))) +
  geom_point() +
  stat_function(fun = function(x) b[1] + b[2]*x, color="red") +
  stat_function(fun = function(x) (b[1] + b[3]) + (b[2] + b[4])*x)
```

When the transmission is 0, it actas as the baseline using the following model:

$$
\hat{Y}_i = \overbrace{13.04}^{\stackrel{\text{y-int}}{\text{baseline}}} + \overbrace{0.3}^{\stackrel{\text{slope}}{\text{baseline}}} X_{1i}
$$

When transmission is 1, we add the difference in y-intercept and slope, which is what $\beta_2$ and $\beta_3$ estimate respectively ($\beta_3$ is known as the "interaction" term), to the baseline y-intercept and slope to get the equation for the second line:

$$
\hat{Y}_i = \underbrace{(\overbrace{13.04}^{\stackrel{\text{y-int}}{\text{baseline}}} - \overbrace{1.36}^{\stackrel{\text{change in}}{\text{y-int}}})}_{\stackrel{\text{y-intercept}}{11.68}} + \underbrace{(\overbrace{.3}^{\stackrel{\text{slope}}{\text{baseline}}} -  \overbrace{0.07}^{\stackrel{\text{change in}}{\text{slope}}})}_{\stackrel{\text{slope}}{.23}} X_{1i}
$$

