---
title: "Expected Move Butterflies"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

## Background

Each week the the "expected move" can be calculated using implied volatility from prices from the options market. The S and P 500 index tends to stay within it's expected move roughly 68 percent of the time. The strategy analyzed here involves buying butterflies at the beginning of each week at the edges of the expected move and selling them at the end of the week, since the edges of the expected move tend to act as magnets and it's very common to see markets pause at those levels. 

I analyzed the past four years worth of weekly expected moves, from 2015 to 2019, to answer the question: is it more profitable to place butterflies each week on the upper end of the expected move, the lower end of the expected move, or on both the upper and the lower?

## Findings

The result of my findings was that for at least the past 4 years, placing butterflies on the upper end of the expected move exclusively is the most profitable strategy. That means if a trader was to place a butterfly only on the upper end every week, they would be more profitable than placing bets on the lower end of the range or placing bets on both the upper and the lower ends.

Unfortunately there doesn't seem to be a big enough difference between low and high volatility* that could be considered statistically significant. There also wasn't a significant interaction between implied volatility and the placement of butterflies, for example neither the upper nor lower end of the range performs significantly differently in either higher or lower volatilties.

*I considered weeks where the expected move was equal to or lower than 3.50 to be low volatility.

## Method

I asked the following questions and hypothesis:

1) Does the placement of the butterfly affect profitability?

Factor: `Placement` with levels $Upper$, $Lower$, and $Combined$.
$$
  H_0: \mu_U = \mu_L = \mu_C = \mu
$$
$$
  H_a: \mu_i \neq \mu \ \text{for at least one}\ i\in\{1=Upper,2=Lower,3=Combined\}
$$

<hr>

2) Does implied volatility affect profitability?

Factor: `Volatility` with levels $High$ and $Low$.
$$
  H_0: \mu_H = \mu_L = \mu
$$
$$
  H_a: \mu_i \neq \mu \ \text{for at least one}\ i\in\{1=H,2=L\}
$$
<hr>

3) Does the effect of Placement change according to different Volatility?

$$
  H_0: \text{The effect of placement is the same for both categories of volatility.}
$$
$$
  H_a: \text{The effect of placement is not the same for both categories of volatility.}
$$
<hr>

$$
\alpha = 0.05
$$

```{r, include=FALSE}
library(mosaic)
library(car)
library(DT)
library(pander)
library(readr)


original_butterflies <- read_csv("../Data/expected-move-butterfly-5.csv")

butterflies <- original_butterflies %>% mutate(Volatility = ifelse(ExpectedMove > 3.5, 'high', 'low'))

upper_result <- butterflies %>% mutate(Volatility = ifelse(ExpectedMove > 3.5, 'high', 'low')) %>% select(c(ClosingDebitUpper, UpperResult, Volatility)) %>% mutate(Placement = 'upper', Profit = UpperResult) %>% select(c(Placement, Profit, Volatility))

lower_result <- butterflies %>% mutate(Volatility = ifelse(ExpectedMove > 3.5, 'high', 'low')) %>% select(c(ClosingDebitLower, LowerResult, Volatility)) %>% mutate(Placement = 'lower', Profit = LowerResult) %>% select(c(Placement, Profit, Volatility))

combined_results <- butterflies %>% mutate(Placement = 'combined') %>% select(c(Placement, CombinedResult, Volatility)) %>% rename(Profit = CombinedResult)

butterflies = c()

butterflies <- rbind(butterflies, combined_results)
butterflies <- rbind(butterflies, upper_result)
butterflies <- rbind(butterflies, lower_result)

butterflies.aov <- aov(Profit ~ Placement + Volatility + Placement:Volatility, data=butterflies)

```

```{r}
pander(summary(butterflies.aov))

xyplot(Profit ~ as.factor(Placement), data=butterflies, type=c("p", "a"), xlab="Placment", main="Significance of placement")

xyplot(Profit ~ as.factor(Volatility), data=butterflies, type=c("p", "a"), xlab="Volatility", main="Non-significance of volatility")

xyplot(Profit ~ as.factor(Placement), data=butterflies, groups=Volatility, type=c("p", "a"), auto.key=list(corner=c(1,1)), xlab="Placement", main="Interaction between volatility and placement")
```

There is some very heavy skew that presents itself in the Q-Q plot. We do have a large enough sample size that this isn't too much of a concern. There may be some concerns with constant variance but I don't believe it's enough to discredit the analysis.

```{r}
par(mfrow=c(1,2))
plot(butterflies.aov, which=1:2, pch=16)
```

### Full Dataset
```{r}
datatable(original_butterflies)
```


