---
title: "Stephanie's Housing Request"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
    toc: true
    toc_float: true
---

**Week 1 Analysis**

```{r, include=FALSE}
library(ggplot2)
library(ggrepel)
library(mosaic)
library(tidyverse)
library(DT) # If you get an error stating: 
            # Error in library(DT): there is no package called 'DT'
            # You will need to run: install.packages("DT") 
            # in your Console, then try "Knit HTML" again.
Rent <- read.csv("../Data/Rent.csv", header=TRUE)
```

<br />

## Background

Stephanie[^1] is a student that will be starting school at BYU-Idaho next semester. Suppose she sent you the following email.

----

"Hi. My name is Stephanie. I would like to learn about what housing options I have for living at BYU-Idaho next semester. It will be my first semester there, so I would like to find something that is close to campus and around $300 a month in rent. I'm not too picky on roomates, but I would like somewhere that has a lot of people around so I can get to know as many people as possible. Thanks in advance!"

----

Write your response to Stephanie below. Use the "Rent" dataset, good statistical analysis, and clear writing to make some well supported suggestions to her about apartments that meet her stated criterions. You are free to use other criterion that you think she might find more meaningful as well.

## Response

Dear Stephanie,
It seems that the criteria you're looking for is something that's cheap and close to campus, and having a lot of people around is a big bonus.

I included a scatterplot of walking distance and the price for each apartment, and you'll notice that the cheapest options and the apartments with the smallest walking distances are on the bottom left of the graph. The three that stand out and that I would recommend looking into first are Normandy Apartments, Bayside Manor, and Birch Plaza.


```{r}
# Code to get you started, be sure to use a subset of Rent instead of Rent in this code though.
datatable(Rent, options=list(lengthMenu = c(3,10,30)), extensions="Responsive")

y_limits <- c(1500, NA)

RentF <- filter(Rent, Gender == "F")
ggplot(RentF, aes(x=Price, y=WalkingMinutes, label = Apartment)) + 
  geom_point(color = ifelse(RentF$Price < 1000 & RentF$WalkingMinutes < 4, "red", "gray")) + 
  geom_label_repel(data = subset(RentF, Price < 1000 & WalkingMinutes < 4,
                  direction     = "x",
                  segment.size  = 0.01,
                  ylim = y_limits,
                  segment.color = "red"
                )) 
```









[^1]: Note that Stephanie is a fictional character who is based on real experiences of many faculty and staff here at BYU-Idaho.
