---
title: "Skills Quiz: Simple Linear Regression"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
    toc: true
    toc_float: true
---


## Instructions

Use this file to keep a record of your work as you complete the "Skills Quiz: Simple Linear Regression" assignment in I-Learn.


----

```{r}
library(pander)
```


<!-- Note: The {} after each Problem and Part header allows you to keep track of what work you have completed. Write something like {Done} once you complete each problem and your html file will then show you a nice summary of what you have "done" already. -->

## Problem 1 {}

Open the `airquality` dataset in R. Perform a regression of daily average `Wind` speed using the daily maximum `Temp`erature as the explanatory variable.

### Part (a) {}

Type out the mathematical equation for this regression model and label both $Y$ and $X$ in the equation.

<div class="YourAnswer">

$$
  \underbrace{Y_i}_\text{Wind} = \beta_0 + \beta_1 \underbrace{\text{X}_i}_\text{Temp} + \epsilon_i \quad \text{where} \ \epsilon_i \sim N(0,\sigma^2)
$$

</div>


### Part (b) {}
 
Fit and summarize a simple linear regression model for this data.

<div class="YourAnswer">

```{r}
someNameYouMake <- lm(Wind ~ Temp, data=airquality)
pander(summary(someNameYouMake))
```

</div>


### Part (c) {}

Type out the estimated equation for this regression model using your estimates found in part (b).

<div class="YourAnswer">

$$
  \hat{Y}_i = 23.234 + -0.170\text{X}_i
$$

</div> 
 
 
### Part (d) {}

Plot the data and the estimated regression model.
 
<div class="YourAnswer">

```{r}
plot(Wind ~ Temp, data=airquality)
abline(someNameYouMake)
```

</div>


### Part (e) {}

Use your model to predict the average daily average Wind speed for an outside temperature of 72 degrees F.

<div class="YourAnswer">

```{r}
pander(predict(someNameYouMake, data.frame(Temp=72), type="response"))
```

</div>


### Part (f) {}

Write out an interpretation of the slope and intercept of your model. Are both meaningful for this data?
 
<div class="YourAnswer">

An increase of 1 degree F in the daily maximum Temperature results in a 0.17 mph decrease in the average daily average Wind speed.

When the daily maximum Temperature is 0 degrees F, the average daily average Wind speed is estimated to be 23.234 mph.

	
Yes, the regression seems to be a good fit to this data and the slope is significant so the interpretation of the slope is meaningful for this data

Yes, the intercept is meaningful for this data because it is significant and an outside temperature of 0 degrees F is a meaningful situation.

</div>



----

## Problem 2 {}

Open the **mtcars** dataset in R. Fit a regression model to the data that can be used to explain average gas mileage of a vehicle (`mpg`) using the weight (`wt`) of the vehicle.

### Part (a) {}

Type out the mathematical equation for this regression model and label both $Y$ and $X$ in the equation.

<div class="YourAnswer">

$$
  \underbrace{Y_i}_\text{mpg} = \beta_0 + \beta_1\underbrace{\text{X}_1}_\text{wt} + \epsilon_i \quad \text{where} \ \epsilon_1 \sim N(0,\sigma^2) 
$$

</div>


### Part (b) {}

Fit and summarize a simple linear regression model for this data.

<div class="YourAnswer">

```{r}
mylm <- lm(mpg ~ wt, data=mtcars)
pander(summary(mylm))
```

</div>


### Part (c) {}

Type out the estimated equation for this regression model using your estimates found in part (b).

<div class="YourAnswer">

$$
  \hat{Y}_i = 37.29 = -5.34\text{X}_i
$$

</div> 
 
 
### Part (d) {}

Plot the data and the estimated regression model.
 
<div class="YourAnswer">

```{r}
plot(mpg ~ wt, data=mtcars)
abline(mylm)
```

</div>


### Part (e) {}

Use your model to predict the average gas mileage (mpg) for a vehicle that weighs 3,000 lbs. (Hint: ?mtcars)

<div class="YourAnswer">

```{r}
predict(mylm, data.frame(wt=3), type="response")
```

</div>


### Part (f) {}

Write out an interpretation of the slope and intercept of your model. Are both meaningful for this data?
 
<div class="YourAnswer">

An increase of 1,000 lbs in the weight of a vehicle results in a 5.34 mpg decrease in the average gas mileage of such vehicles.

	
The average gas mileage of vehicles that weigh nothing (0 lbs) is estimated to be 37.29 mpg.

While the slope is significant, it only looks to be meaningful for vehicles that weigh between 2.5 thousand and 4 thousand pounds. Otherwise this regression does not seem to be a good fit for this data based on the scatterplot.

No, the intercept is not meaningful for this data because a vehicle with weight zero is not possible.

</div>



----

Before we can really trust the interpretation of and predictions from a regression model, there are important diagnostic checks to perform on the regression. These diagnostics are even more important to perform when p-values or confidence intervals are used as part of the analysis. In future weeks of this course, we will focus in greater detail on the technical details of regression: hypothesis tests, confidence intervals, and diagnostic checks. However, for the sake of completeness, the following problems have run through these technical details, even though we lack full understanding about them for the time being.



----

## Problem 3 {}

Use your regression for the `airquality` data set in **Problem 1** to complete the following "technical details" for this regression.

### Part (a) {}

Create a (1) residuals vs. fitted-values plot, (2) Q-Q Plot of the residuals, and (3) residuals vs. order plot.

<div class="YourAnswer">

```{r}
par(mfrow=c(1,3))
plot(mylm, which=1:2)
plot(mylm$residuals, main="Residuals vs Order", xlab="", ylab="Residuals")
```

</div>


### Part (b) {}

Explain, as best you understand currently, what each of these three plots show for this regression.

<div class="YourAnswer">

Everything looks pretty good. The residuals vs. fitted-values plot shows constant variance and a nice linear relation. The Q-Q Plot shows possible problems with normality because some dots go out of bounds, but is fairly good. The residuals vs. order plot shows no problems with time trends, so the error terms can be assumed to be independent.

</div>


### Part (c) {}

Report the p-value for the test of these hypotheses for your regression.

**Intercept Hypotheses**

$$
  H_0: \beta_0 = 0 \quad \text{vs.} \quad H_a: \beta_0 \neq 0
$$

< 2e-16

**Slope Hypotheses**

$$
  H_0: \beta_1 = 0 \quad \text{vs.} \quad  H_a: \beta_1 \neq 0
$$
< 2.64e-09


Comment on whether or not we should trust each p-value based on your plots in Part (a). 

<div class="YourAnswer">

Yes, they can be trusted because the diagnostic plots all checked out

</div>



----

## Problem 4 {}

Use your regression for the `mtcars` data set in **Problem 2** to complete the following "technical details" for this regression.

### Part (a) {}

Create a (1) residuals vs. fitted-values plot, (2) Q-Q Plot of the residuals, and (3) residuals vs. order plot.

<div class="YourAnswer">

```{r}
par(mfrow=c(1,3))
plot(someNameYouMake, which=1:2)
plot(someNameYouMake$residuals, main="Residuals vs Order", xlab="", ylab="Residuals")
```

</div>


### Part (b) {}

Explain, as best you understand currently, what each of these three plots show for this regression.

<div class="YourAnswer">

Everything looks somewhat questionable. The residuals vs. fitted-values plot shows a lack of linearity, which makes it hard to judge constant variance. The Q-Q Plot shows possible problems with normality because some dots go out of bounds, but is fairly good. The residuals vs. order plot shows a possible problem with time trends due to the slight rainbow pattern.

</div>


### Part (c) {}

Report the p-value for the test of these hypotheses for your regression.

**Intercept Hypotheses**

$$
  H_0: \beta_0 = 0 \quad \text{vs.} \quad H_a: \beta_0 \neq 0
$$
< 2e-16

**Slope Hypotheses**

$$
  H_0: \beta_1 = 0 \quad \text{vs.} \quad  H_a: \beta_1 \neq 0
$$

1.29e-10

Comment on whether or not we should trust each p-value based on your plots in Part (a). 

<div class="YourAnswer">

No, it should not be trusted because of the lack of linearity and the distance zero is from the current date.

</div>













<style>

.YourAnswer {
  color: #317eac;
  padding: 10px;
  border-style: solid;
  border-width: 2px;
  border-color: skyblue4;
  border-radius: 5px;
}

</style>

 
 